### 2016-06-26 ch02-qlogms/excitation-design.tex
Small corrections CH02

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. 
Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not know at the time of the design.
There are two main solutions to this in the engineering literature~\citep{Goodwin2006GBO}:
\begin{enumerate}
\item Iterative input design where the design updated as more information about the system is available, see e.g.~\citep{Hjalmarsson2005,Gevers2005}.
The disadvantage of this approach is that multiple experiments are required to converge to the optimal design.
Moreover, since the signal is optimized for a specific system model, it is likely that changes in the dynamics are hard to observe using such a signal.

@@

\begin{remark}
D-optimal input designs are quite popular since this criterion minimizes the volume of uncertainty ellipsoid of the parameters.
Moreover, the D-optimal input design is independent of the parametrization of the system (as long as the different parametrizations can be related by a bijective function)~\citep[Chapter 6]{Goodwin1977}.
\end{remark}

@@

Furthermore, when the damping of the systems is low or the resonances of the sub-systems are well separated, the estimates decouple in the sense that the elements off-diagonal blocks become small.
The effect is that the variance of the estimated transfer function approximately becomes the sum of the variances of the sub-systems.
To illustrate this we look at a system with $K=2$, i.e. the sum of two second order systems.
The conclusions from the example are easily extended to systems consisting of more than two sub-systems.

@@

\begin{assumption}
The resonance frequencies $\wn[i]$ and $\wn[j]$ of all the different sub-systems in \eqref{eq:excitation:PFE} are well-separated in the frequency domain.
\end{assumption}

Under these assumptions, the Fisher information matrix will be approximately a diagonal block matrix, where the $i^{\text{th}}$ block corresponds to the parameters $\left[\wn[i], \wz[i], \damping[i], \dc[i,1], \ldots, \dc[i,M_{2,i}] \right]$ of a particular sub-system.
Using the same logic as in \secref{sec:excitation:ex:sumOfSecondOrderSipmle}, this means that the variance of the complete system is approximately equal to the variance obtained for each sub-system, due to the decoupling of the Fisher information matrix.

@@

  When designing an excitation signal, it is common only to specify a (large) frequency band by its boundary frequencies $f_{\min}$ and $f_{\max}$.
  For a given $\alpha$ it is easy to determine the number of excited lines within the frequency range of interest:
    \begin{equation}
      F = \floor{
                  \frac{\log{f_{\max}} - \log{f_{\min}}}
                       {\log{\alpha}}
                }
    \label{eq:excitation:logms:F}
    \end{equation}
  where $\floor{\bullet}$ denotes rounding towards the nearest smaller integer value.
  In the Section \ref{sec:excitation:optimAlpha}, the choice of a suitable $\alpha$ is elaborated.

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternate specifications are equivalent.
  \end{remark}
  \begin{example} \label{eg:excitation:logarithmicGrid:matlab}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  This allows to define the amplitude spectrum $A_k$ as
  \begin{equation}
    A_k = A_{\mathrm{in}}
                 \sqrt{\int_{\fMin{k}}
                     ^{\fMax{k}}
                     \limits
                     \frac{1}
                          {f}
                     \,\dd{f}}
          = A_{\mathrm{in}}
                \cdot
                \sqrt{
                \ln \frac{\;\fMax{k}\;}{\fMin{k}}
                }
  \end{equation}
  where $\left[ \fMin{k}, \fMax{k} \right] $ denote the frequency range over which \gls{PSD} of the pink noise is to be approximated by the power at frequency line $f_k$.

@@

  \begin{figure}[ht]
    \centering
      \setlength\figurewidth{0.8\columnwidth}
      \setlength\figureheight{0.5\figurewidth}
    \input{\thisDir/figs/freqGrid.tikz}
    \caption[Amplitude spectrum for linear, logarithmic, quasi-logarithmic grids.]{Comparison of the amplitude spectrum of a linear grid multisine, a logarithmic grid multisine, the compensated quasi-logarithmic multisine and the `pink' quasi-log multisine.
    $\omegaLinLog$ indicates the cross-over between the linear and logarithmic regions. }%
    \label{fig:excitation:freqGrids}
  \end{figure}

@@

  We can expect a very similar behavior from properly designed compensated quasi-log multisines and from the `pink' quasi-log multisine, since their amplitude spectra are almost identical.
  As such, it is advised to use `pink' quasi-log multisines as those can be constructed using more straightforward code (equation \eqref{eq:excitation:pink}) than the compensated quasi-log multisines (equations \eqref{eq:excitation:nfk}, \eqref{eq:excitation:nearestLines} , and \eqref{eq:excitation:compensationAk}).

@@

\begin{remark}
\citet{Goodwin1977} note that for optimal input designs:
\begin{quote}
The choice of criteria is often not critical since it is usually the case that a good experiment according to one criterion will be deemed good by other criteria. 
Naturally this will only be true for sensibly chosen criteria.
\end{quote}
Also for the design of robust excitation signals, this is worthwhile observation.
\end{remark}

@@

  We then calculate the theoretical variance on the magnitude of the peak in the \gls{FRF} ($\abs{ G\left( \wM \right) }$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.

@@

  We do see, however, that the variance in the vicinity of the resonance of every system is lower than in the case of a linearly spaced excitation.

@@

We observe that the amplitude of the transfer function is $9.8\unit{dB}$ and $8.1\unit{dB}$ at the resonant frequencies.
The standard deviation on the \glspl{FRF} lies respectively $59.7 \unit{dB}$ and $59.6 \unit{dB}$ lower than the \gls{FRF} at these frequencies. 
This shows that both resonances are captured with a practically identical relative error.

@@

In this chapter we have illustrated that quasi-logarithmic multisine excitations are a well-suited to measure transfer functions over broad frequency ranges with a constant relative uncertainty of the present resonances.
The suggested amplitude compensation methods allow the user to measure a parametric \gls{LTI} model for systems in an even wider frequency band with a constant uncertainty.
The variance of the estimates at lower frequencies is improved in comparison to the uncompensated case at the cost of a slightly increased (yet constant) variance at higher frequencies for excitation signals with an equal power content.
In practice, this slight increase is likely to be negligible.
The suggested compensation works well as long as the considered systems are sufficiently damped with respect to the density of the excited frequency lines.

>>>>

A complicating issue in identification of dynamic system is the fact that the information matrix often depends on the true parameter values. 
Consequently, an optimal input design, based on the information matrix, will depend on quantities that are not known at the time of the design.
There are two main solutions to this in the engineering literature~\citep{Goodwin2006GBO}:
\begin{enumerate}
\item Iterative input design where the design is updated as more information about the system is available, see e.g.~\citep{Hjalmarsson2005,Gevers2005}.
The disadvantage of this approach is that multiple experiments are required to converge to the optimal design.
Moreover, since the signal is optimized for a specific system model, it is likely that changes in the dynamics are hard to observe using such a signal.

@@

\begin{remark}
D-optimal input designs are quite popular since this criterion minimizes the volume of the uncertainty ellipsoid of the parameters.
Moreover, the D-optimal input design is independent of the parametrization of the system (as long as the different parametrizations can be related by a bijective function)~\citep[Chapter 6]{Goodwin1977}.
\end{remark}

@@

Furthermore, when the damping of the systems is low or the resonances of the sub-systems are well separated, the estimates decouple in the sense that the elements of the off-diagonal blocks become small.
The effect is that the variance of the estimated transfer function approximately becomes the sum of the variances of the sub-systems.
To illustrate this we look at a system with $K=2$, i.e. the sum of two second order systems.
The conclusions from the example are easily extended to systems consisting of more than two sub-systems.

@@

\begin{assumption}
The resonance frequencies $\wn[i]$ and $\wn[j]$ of all the different sub-systems in \eqref{eq:excitation:PFE} are well-separated in the frequency domain. %TODO: how well-separated?
\end{assumption}

Under these assumptions, the Fisher information matrix will be approximately a diagonal block matrix, where the $i^{\text{th}}$ block corresponds to the parameters $\left[\wn[i], \wz[i], \damping[i], \dc[i,1], \ldots, \dc[i,M_{2,i}] \right]$ of a particular sub-system.
Using the same logic as in \secref{sec:excitation:ex:sumOfSecondOrderSipmle}, this means that the variance of the complete system is approximately equal to the sum of the variances obtained for each sub-system, due to the decoupling of the Fisher information matrix.

@@

  When designing an excitation signal, it is common to only specify a (large) frequency band by its boundary frequencies $f_{\min}$ and $f_{\max}$.
  For a given $\alpha$ it is easy to determine the number of excited lines within the frequency range of interest:
    \begin{equation}
      F = \floor{
                  \frac{\log{f_{\max}} - \log{f_{\min}}}
                       {\log{\alpha}}
                }
    \label{eq:excitation:logms:F}
    \end{equation}
  where $\floor{\bullet}$ denotes rounding towards the nearest smaller integer value.
  In the Section \ref{sec:excitation:optimAlpha}, the choice of a suitable $\alpha$ is elaborated.

  \begin{property}\label{prop:excitation:log:relBW}
   Each frequency band $B = \ClosedInterval{f, \kappa f}$ with $\kappa \in \PositiveRealNumbers \without \set{0,1}$ that is completely within the range of the excited logarithmic grid of frequencies $\set{f_1, \ldots, f_F}$ contains a constant number of frequency lines
   \begin{equation}
     F_{B} \isdef
     \floor{\frac{\log (\kappa f) - \log( f )}{\log \alpha}}
         = \floor{\log_{\alpha} \kappa}
         \text{.}
   \end{equation}
   \end{property}
  \begin{example}
    For the frequency decade band $\decade(f) \isdef \ClosedInterval{f, 10 f}$, a logarithmic frequency grid contains $F_{\decade} = \floor{\log_{\alpha} 10}$ excited lines.
    Note that this is independent of the frequency $f$.
  \end{example}
  \begin{remark}
    In different contexts, e.g. electrical circuit simulators such as \gls{SPICE}~\citep{Kundert1995}, measurement equipment such as \glspl{DSA} and programming languages, it is more common to specify the spacing of a logarithmic grid using the total number of frequencies ($F$) or the number of frequencies per decade ($F_{\decade}$) instead of the grid ratio $\alpha$.
    Obviously, such alternate specifications are equivalent.
  \end{remark}
  \begin{example} \label{eg:excitation:logarithmicGrid:matlab}
    The \MATLAB code \mcode[mathescape]{f = logspace(log10($f_{\min}$), log10($f_{\max}$), $F$)} produces a logarithmically spaced frequency vector \mcode{f} with $F$ elements in the range $\left[ f_{\min}, f_{\max} \right]$.
    As such, the effective frequency ratio $\alpha = \sqrt[F]{\frac{f_{\max}}{f_{\min}}}$ in this vector.
  \end{example}
  \begin{example}
    The~\citet{HP3562A} has a logarithmic measurement mode that ensures that each decade is covered by $F_{\decade} = 80$ lines~ to measure an \gls{FRF}.
    %TODO: make sure "\gls{FRF}" has already been used before this point
    This is equivalent to a grid spacing $\alpha = \sqrt[80]{10} \approx 1.029$.
  \end{example}

@@

  This allows to define the amplitude spectrum $A_k$ as
  \begin{equation}
    A_k = A_{\mathrm{in}}
                 \sqrt{\int_{\fMin{k}}
                     ^{\fMax{k}}
                     \limits
                     \frac{1}
                          {f}
                     \,\dd{f}}
          = A_{\mathrm{in}}
                \cdot
                \sqrt{
                \ln \frac{\;\fMax{k}\;}{\fMin{k}}
                }
    \label{eq:excitation:pinkMS:Ak}
  \end{equation}
  where $\left[ \,\fMin{k}, \fMax{k}\, \right] $ denotes the frequency range over which the \gls{PSD} of pink noise is to be approximated by the power at frequency line $f_k$.

@@

  \begin{figure}[ht]
    \centering
      \setlength\figurewidth{0.8\columnwidth}
      \setlength\figureheight{0.5\figurewidth}
    \input{\thisDir/figs/freqGrid.tikz}
    \caption[Amplitude spectrum for linear, logarithmic, quasi-logarithmic grids.]{Comparison of the amplitude spectrum of a linear grid multisine, a logarithmic grid multisine, the compensated quasi-logarithmic multisine and the `pink' quasi-log multisine covering the frequency band $\ClosedInterval{1\unit{Hz}, 1 \unit{kHz}}$.
    $\omegaLinLog$ indicates the cross-over between the linear and logarithmic regions in the lower two decades of the excited frequency band.}%
    \label{fig:excitation:freqGrids}
  \end{figure}

@@

  We can expect a very similar behavior from properly designed compensated quasi-log multisines and from the `pink' quasi-log multisine, since their amplitude spectra are almost identical.
  As such, it is advised to use `pink' quasi-log multisines as those can be constructed using more straightforward code (equation \eqref{eq:excitation:pinkMS:Ak}) than the compensated quasi-log multisines (equations \eqref{eq:excitation:nfk}, \eqref{eq:excitation:nearestLines} , and \eqref{eq:excitation:compensationAk}).

@@

\begin{remark}
\citet{Goodwin1977} note that for optimal input designs:
\begin{quote}
The choice of criteria is often not critical since it is usually the case that a good experiment according to one criterion will be deemed good by other criteria. 
Naturally this will only be true for sensibly chosen criteria.
\end{quote}
Also for the design of robust excitation signals, this is a worthwhile observation.
\end{remark}

@@

  We then calculate the theoretical variance on the magnitude of the peak in the \gls{FRF} ($\abs{ G\left( \wM \right) }$) as shown in \eqref{eq:excitation:varPeak} for a coarse grid.
  %TODO: clarify previous sentence
  A coarse grid was chosen with $\tilde{\alpha} = 10$; this means that the spacing between the excited lines around the resonance are much larger than the system bandwidth.
  It is obvious that such an excitation signal will not perform well in most cases, as the resonance peak is very likely to be overlooked in such an \gls{FRF}.

@@

  We do see, however, that the variance in the vicinity of the resonance of most systems is lower than in the case of a linearly spaced excitation.

@@

We observe that the amplitude of the transfer function is $9.8\unit{dB}$ and $8.1\unit{dB}$ at the resonance frequencies.
The standard deviation on the \glspl{FRF} lies respectively $59.7 \unit{dB}$ and $59.6 \unit{dB}$ lower than the \gls{FRF} at these frequencies. 
This shows that both resonances are captured with a practically identical relative error.

@@

In this chapter we have illustrated that quasi-logarithmic multisine excitations are well-suited to measure transfer functions over broad frequency ranges with a constant relative uncertainty of the present resonances.
The suggested amplitude compensation methods allow the user to measure a parametric \gls{LTI} model for systems in an even wider frequency band with a constant uncertainty.
The variance of the estimates at lower frequencies is improved in comparison to the uncompensated case at the cost of a slightly increased (yet constant) variance at higher frequencies for excitation signals with an equal power content.
In practice, this slight increase is likely to be negligible.
The suggested compensation works well as long as the considered systems are sufficiently damped with respect to the density of the excited frequency lines.
